---
title: PA1_template
output: md_document
---
Project 1 Assignment - Reproducible Research
============================================

Here, we're going to load the activity data for the assignment:

```{r}
#Call to load the dataset:
activity <- read.csv("C:/Users/Phil/Documents/Data_Science/Reproducible_Research/activity.csv")

#Call to omit all NA values from the dataset
activity_naremoved <- na.omit(activity)

#Call to summarize the modified dataset
summary(activity_naremoved)
```

Next, we're going to calculate a few items from the dataset:

First, we'll calculate the total number of steps taken per day. To do this, we'll use the aggregate function 

```{r}
#Calculates the total number of steps taken per day
number_steps_per_date <- aggregate(activity_naremoved$steps, list(activity_naremoved$date), sum)
colnames(number_steps_per_date) <- c("date", "steps")
```

Next, we'll plot a histogram using this data
```{r}
#Plots a histogram of the data using library ggplot2
library(ggplot2)
ggplot(data=number_steps_per_date, aes(x=steps)) +
        geom_histogram() +
        ggtitle("Total Steps Taken Per Day") +
        labs(x="Number of Steps per Day", y="Numbrt of Times per Day")
```

We now want to calculate the mean and median total steps per date from our dataset:

```{r}
#Calculate the mean and median
mean_steps_per_date <- mean(number_steps_per_date$steps)
median_steps_per_date <- median(number_steps_per_date$steps)

#Outputs the mean and median
mean_steps_per_date
median_steps_per_date
```

```{r}
#Time Series Plot
steps_per_interval <- aggregate(activity_naremoved$steps, by = list(interval = as.factor(activity_naremoved$interval)), FUN=mean, na.rm=TRUE)
steps_per_interval$interval <- as.integer(levels(steps_per_interval$interval)[steps_per_interval$interval])
colnames(steps_per_interval) <- c("interval", "steps")

time_series <- tapply(activity_naremoved$steps, activity_naremoved$interval, mean, na.rm = TRUE)

plot(row.names(time_series), time_series, type = "l", xlab = "5-Min Interval", 
    ylab = "Average Over All Days", main = "Average Number of Steps Taken", 
    col ="gray")

#Calculation of Max Interval
max_interval <- steps_per_interval[which.max(steps_per_interval$steps),]
max_interval
```

``` {r}
#Calculates and returns the total count of NA values in the dataset
total_na <- sum(is.na(activity))
total_na

#Aggregates a simplified dataset, similar to above
number_steps_per_date_wna <- aggregate(activity$steps, list(activity$date), sum)
colnames(number_steps_per_date_wna) <- c("date", "steps")

#A loop that runs through the steps column of the dataset, inputting the mean steps per day where NA is present to create a new dataset
for (i in 1:nrow(number_steps_per_date_wna)){
        tmp <- number_steps_per_date_wna$steps[i]
        if(is.na(tmp)){
                number_steps_per_date_wna$steps[i] <- mean_steps_per_date
        }
}
#Prints the header of the new dataset
head(number_steps_per_date_wna)

#Plots the new dataset
ggplot(data=number_steps_per_date_wna, aes(x=steps)) +
        geom_histogram() +
        ggtitle("Total Steps Taken Per Day") +
        labs(x="Number of Steps per Day", y="Numbrt of Times per Day")

#Calculation of new mean and median
new_mean <- mean(number_steps_per_date_wna$steps)
new_median <- median(number_steps_per_date_wna$steps)

new_mean
new_median

#The new value of the median differs slightly from the previous calculation, while the means have remained the same.
#The result of imputing missing data on the estimates is very small, and arguably insigificant
```     

```{r}
#Code to convert dates into Weekend or Weekdays, using as.Date and a for loop, with if call
date_convert <- as.Date(activity_naremoved$date)
weekdays_dataset <- weekdays(date_convert, abbreviate = TRUE)

for (i in 1:length(weekdays_dataset)) {
        tmp2 <- weekdays_dataset[i] 
        if (tmp2 == "Sat") {
                weekdays_dataset[i] <- "Weekend"
        }
                else if (tmp2 == "Sun") { 
                        weekdays_dataset[i] <- "Weekend"        
                }
                else {
                        weekdays_dataset[i] <- "Weekday"  
                }
}

#Replace dates with Weekday or Weekend
activity_naremoved_2 <- activity_naremoved
activity_naremoved_2$date <- weekdays_dataset
activity_naremoved_2$date <- factor(activity_naremoved_2$date)

steps_per_day_2 <- aggregate(steps ~ interval + weekdays_dataset, data = activity_naremoved, FUN ="mean")
names(steps_per_day_2) <- c("interval", "date", "steps")

#Time Series Plot #2
library(lattice)
xyplot(steps_per_day_2$steps ~ steps_per_day_2$interval | steps_per_day_2$date, layout = c(1, 2), type = "l", xlab = "Interval", ylab = "Number of Steps")

